\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}

### 2022-01-17 sections/methods.tex
Change: Update consensus algorithm subsection in the Methods section

  \subsubsection*{Bootstrapping}
  For each network inference method (both correlation and direct association based methods), $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
  We then recalculate the associations in these permuted \ac{otu} tables using the different network inference algorithms.
  Finally, we calculate the p-value based on how often a more extreme association is observed for randomly permuted data.

  \subsubsection*{p-value merging (can be moved to the supplementary section)}
  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $P_i$, the statistic $\Psi$:
  \begin{equation*}
    \begin{aligned}
        \Psi &= \sum_{i=1}^k -2 \log \left( P_i \right) \\
        \Psi &\sim \chi^2_{2k}
    \end{aligned}
  \end{equation*}

  Brown~\cite{brown_400_1975} extended Fisher's method to dependent p-values by using a re-scaled $\chi^2$ distribution:
  \begin{equation*}
    \Psi \sim c \chi^2_{2f}
  \end{equation*}
  where, $f$ is the degrees of freedom and $c$ is the scale factor and are given by:
  \begin{equation*}
    f = \frac{\mathrm{E}[\Psi]^2}{\mathrm{Var}[\Psi]} ~~~\text{and}~~~ c = \frac{\mathrm{Var}[\Psi]}{2\mathrm{E}[\Psi]} = \frac{k}{f}
  \end{equation*}

  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated directly via a numerical integration:
  \begin{equation*}
    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right)
  \end{equation*}

  Kost and McDermott~\cite{kost_combining_2002} further fit a third-order polynomial to approximate the covariance
  \begin{equation}
    \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
    \label{eqn:covariance-pvalues}
  \end{equation}
  where, $\rho_{ij}$ is the correlation between method $i$ and method $j$

  The final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
  \begin{equation}
    \begin{aligned}
        & P_{combined} = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(P_i) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
    \end{aligned}
    \label{eqn:pvalue-combined}
  \end{equation}

  The p-value merging method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
  % FIXME: Not sure what our stance on this is?
  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
  The combined p-values are used to threshold for significance during the consensus network step.

 \subsection*{Notation}

@@

  $N^i$ = The $i^{th}$ occurrence network represented as a $q \times q$ adjacency matrix, where $i \in {1, q}$

@@

  $\bar{N}^i$ is a "flattened" version of the adjacency matrix $N^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector:

  \subsection*{Consensus method}
  \vspace{-5mm}

@@

  \subsection*{Algorithm 1: Simple voting}
   \vspace{-5mm}

@@

  \subsubsection*{Algorithm 2: Scaled-sum method}
   \vspace{-5mm}

@@

  \begin{equation}
    S^{i} = \frac{N^{i}}{\max(\mid N^{i} \mid)}
    \label{eqn:scaled-sum-rescaling}
  \end{equation}

>>>>

 \subsubsection*{Notation}

@@

  $N^i$ = The weights of the $i^{th}$ co-occurrence network represented as a $q \times q$ adjacency matrix, where $i \in {1, q}$
  $P^i$ = The pvalues of the $i^{th}$ co-occurrence network represented as a $q \times q$ adjacency matrix, where $i \in {1, q}$

@@

  $\bar{N}^i$ is a "flattened" version of the adjacency matrix $N^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector:
  $\bar{P}^i$ is a "flattened" version of the adjacency matrix $P^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector:

  \subsubsection*{Bootstrapping and p-value calculation}
  For all correlation-based methods $k \in 1:w$, $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
  All the direct association based methods used in the study have their own regularization methods built-in and hence do not need to undergo bootstrapping.
  We then recalculate the correlations in these permuted \ac{otu} tables using the different correlation-based algorithms.
  Finally, we calculate the p-value based on how often a more extreme association is observed for randomly permuted data.

  \subsubsection*{p-value merging}

  The first part of the consensus algorithm workflow is to merge the p-values for the networks generated by the correlation-based methods.
  This step is performed using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016} (refer Supplementary).

  % TODO: Is this necessary here, should this be elaborated?
  The final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
  \begin{equation}
    \begin{aligned}
        & \hat{P}_j = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
    \end{aligned}
    \label{eqn:pvalue-combined}
  \end{equation}
  where, $\hat{P}_j$ is the combined p-value for the edge $j$

  % FIXME: Not sure what our stance on this is?
  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
  The combined p-values are used to threshold for significance right before the consensus algorithm is applied to the inferred networks.

  \subsubsection*{Consensus method}
  \vspace{-5mm}

@@

  \paragraph*{Algorithm 1: Simple voting}

@@

  \paragraph*{Algorithm 2: Scaled-sum method}

@@

  \begin{equation}
    S^{i} = \frac{\bar{N}^{i}}{\max(\mid \bar{N}^{i} \mid)}
    \label{eqn:scaled-sum-rescaling}
  \end{equation}

### 2022-01-17 sections/supplementary.tex
Change: Update consensus algorithm subsection in the Methods section

>>>>

  \subsection{Consensus and p-value}

  % FIXME: Should this section be rewritten? Ask Daniel
  \subsubsection*{p-value merging}
  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $P_i$, the statistic $\Psi$:
  \begin{equation*}
    \begin{aligned}
        \Psi &= \sum_{i=1}^k -2 \log \left( P_i \right) \\
        \Psi &\sim \chi^2_{2k}
    \end{aligned}
  \end{equation*}

  Brown~\cite{brown_400_1975} extended Fisher's method to dependent p-values by using a re-scaled $\chi^2$ distribution:
  \begin{equation*}
    \Psi \sim c \chi^2_{2f}
  \end{equation*}
  where, $f$ is the degrees of freedom and $c$ is the scale factor and are given by:
  \begin{equation*}
    f = \frac{\mathrm{E}[\Psi]^2}{\mathrm{Var}[\Psi]} ~~~\text{and}~~~ c = \frac{\mathrm{Var}[\Psi]}{2\mathrm{E}[\Psi]} = \frac{k}{f}
  \end{equation*}

  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated directly via a numerical integration:
  \begin{equation*}
    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right)
  \end{equation*}

  Kost and McDermott~\cite{kost_combining_2002} further fit a third-order polynomial to approximate the covariance
  \begin{equation}
    \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
    \label{eqn:covariance-pvalues}
  \end{equation}
  where, $\rho_{ij}$ is the correlation between method $i$ and method $j$

  The final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
  \begin{equation}
    \begin{aligned}
        & P_{combined} = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(P_i) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
    \end{aligned}
    \label{eqn:pvalue-combined}
  \end{equation}

  The p-value merging method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
  % FIXME: Not sure what our stance on this is?
  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
  The combined p-values are used to threshold for significance during the consensus network step.
